---
title: "Data Exploration: AI_Combined_File_Final"
author: "Masha Logan"
date: "3/17/2020"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Description: 

This part describe the overall quality and important findings for each data set used in the analysis.  This is equivalent an “Executive Summary” of a regular report


```{r, echo = FALSE, warning = FALSE, message = FALSE}

.libPaths("C:/Users/mlogan/Desktop/R_Packages")

library(tidyverse)
library(ggstance)
library(patchwork)
library(ggExtra)
library(e1071) ##for skewness
library(lubridate) ## for checking the dates
library(knitr)
library(kableExtra) ## for pretty tables
library(stats) ## for setting names
library(crayon)

```


```{r, warning = FALSE, message = FALSE, echo = FALSE}
## read the file in 

sample_data<-read_csv("C:/Users/mlogan/Desktop/AI_Combined_File_Final.csv", 
         guess_max = 264164, na = c("N/A", "", "NA", "NULL", "<NA>", "-"))

## rename the file
dr_data <-as.data.frame(sample_data)

```

The dataset contains `r nrow(sample_data)` observations of `r ncol(sample_data)` variables.

```{r, echo = FALSE}
### convert characters to factors

dt<-sapply(dr_data, typeof) ## extract types of data from our sample
char_names<-names(which(dt == "character")) ## see what are characters

## also see that everything that starts with "src_" gets converted to factor as well

src_names<-grep("^src", names(dr_data), value=TRUE)
no_names<-grep("^no_of_", names(dr_data), value=TRUE)

## convert all char to factors
##dr_data[char_names]<-lapply(dr_data[char_names], factor)
dr_data[src_names]<-lapply(dr_data[src_names], as.character)


## let's see what data types we have here

var_types<-sapply(dr_data, class)
var_types<-as.data.frame(var_types)

# make rownames to be the first column 

var_types<-rownames_to_column(var_types)
var_types<-var_types%>%arrange(var_types)

```

The types of these variables are the following:

```{r, echo = FALSE}

## see variable types in the dataset

var_types_summary<-var_types%>%group_by(var_types)%>%count()%>%arrange(n) %>%ungroup

var_types_summary%>%kable(col.names = c('Data Type', '# of Variables')) %>%
    kable_styling("bordered", full_width = F, position = "left")

```


```{r, echo = FALSE}
## let's subset our dataset by var type

dr_data_num_cont<- dr_data%>%select_if(is.numeric)%>%select(-no_names, -MOD, -CLAIMCOUNT, LITINDICATOR) ## all contd vars
dr_data_num_desc<- dr_data%>%select(no_names, MOD, CLAIMCOUNT, LITINDICATOR) ## discrete vars
dr_data_fact<- dr_data%>%select_if(is.factor)
dr_data_date<- dr_data%>%select_if(is.Date)
dr_data_char<- dr_data%>%select_if(is.character)

```


### Description of each variable

</p>

#### Timestamp variables
</p>


```{r timestamp, echo = FALSE, message = FALSE, warning = FALSE, fig.width= 6, fig.height=3.5, fig.align = "left", comment = NA}


for(i in names(dr_data_date)) {
  
## print the variable name
  
  cat("-", paste(i), "\n")
  cat("\n")
  
 ## 5 number summary 
  print(summary(dr_data_date[, i])) 
  
  cat("\n")
  cat("\n")
  
}


```


#### Categorical Variables
</p>


```{r categorical, echo = FALSE, message = FALSE, warning = FALSE, fig.width= 6, fig.height=3.5, fig.align = "left", comment = NA}


for(i in names(dr_data_char)) {
  
## print the variable name
  
  cat("-", paste(i), "\n")
  cat("\n")
  
  
## bar chart

# p3<-ggplot(dr_data_char, aes(x = dr_data_char[,i])) +
#   geom_bar(stat= 'count', color = "black", fill = "lightgray") + 
#   labs (title = paste(i),
#         x = paste (i),
#         y = '') +
#   theme_bw() +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1))
# 
# print(p3)
  

## Cardinality

cardinal<-dr_data_char[,i] %>% n_distinct()
cat(paste("Cardinality:", cardinal, "\n"))

## Missing values

missing<- sum(is.na(dr_data_char[,i]))
cat(paste("Missing Values:", missing, "\n"))

## Smallest Cell Value and Frequency

small<-dr_data_char %>% count(dr_data_char[,i])%>%arrange(n)%>%filter(n == min(n))%>%
setNames(c('x','n'))
cat("Smallest Cell Value and Frequency:", "\n")
## table

print(small)
cat("\n")
cat("\n")

}

```


#### Numeric Variables
</p>

```{r numeric_cont, echo = FALSE, message = FALSE, warning = FALSE, fig.width= 4, fig.height=3.5, fig.align = "left", comment = NA}

## numeric variables - continuous only


sub_data<-dr_data_num_cont

for(i in names(sub_data)) {
  
  ## density plot
  p1<-sub_data %>% ggplot() + 
    ## geom_histogram(aes(x = sub_data[,i])) +
    geom_density(aes(x = sub_data[,i])) +
    labs (title = paste(i),
          x = '',
          y = '') +
    theme_bw()
  
  ## boxplot
  p2<- sub_data %>% 
    ## geom_histogram(aes(x = sub_data[,i])) +
    ggplot(aes(y = '', x = sub_data[,i])) +
    geom_boxploth(outlier.colour = "red") +
    labs (x = paste (i),
          y = '') +
    theme_bw()
  
  ## gotta use print to show the plot
  print(p1/p2)
  
  ## 5 number summary 
  print(summary(sub_data[, i])) 
  
  ## Skewness rounded to 4 decimal places
  
  cat(paste("The skewness is", round(skewness(sub_data[, i], na.rm =T),4)), "\n")
  
  ## show the top 5 smallest values 
  cat("Five smallest values are: ") 
  cat(head(sort(sub_data[, i], decreasing=F), 5),"\n" )
  
  ## show the top largest values
  cat("Five largest values are: ") 
  cat(head(sort(sub_data[, i], decreasing = T),5), "\n")
  
}
```



```{r numeric_desc, echo = FALSE, message = FALSE, warning = FALSE, fig.width= 4, fig.height=3.5, fig.align = "left", comment = NA}

## numeric variables - continuous only


sub_data<-dr_data_num_desc

for(i in names(sub_data)) {
  
  ## barchart plot

  p3<-ggplot(sub_data, aes(x = sub_data[,i])) +
  geom_bar(stat= 'count', color = "black", fill = "lightgray") +
  labs (title = paste(i),
        x = paste (i),
        y = '') +
        theme_bw()

  print(p3)
  
  
  ## 5 number summary 
  print(summary(sub_data[, i])) 
  
  ## Skewness rounded to 4 decimal places
  
  cat(paste("The skewness is", round(skewness(sub_data[, i], na.rm =T),4)), "\n")
  
  ## show the top 5 smallest values 
  cat("Five smallest values are: ") 
  cat(head(sort(sub_data[, i], decreasing=F), 5),"\n" )
  
  ## show the top largest values
  cat("Five largest values are: ") 
  cat(head(sort(sub_data[, i], decreasing = T),5), "\n")
  
}
```
